---
title: "Examples for the corrgram package"
author: "Kevin Wright"
date: "`r Sys.Date()`"
bibliography: corrgram.bib
output:
  html_document:
    highlight: tango
vignette: >
  %\VignetteIndexEntry{Examples for the corrgram package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
<style type="text/css">
body, td { font-size: 14pt; color: #000; padding:1em; }
code, pre, #ws, #message { font-size: 12pt;
  color: #000; background-color: #f5f5f5; }
code {border: 1px solid #e0e0e0; padding: 0 5px;}
pre > code { border: 0; }
</style>

## Abstract

The `corrgram` package is an implementation of correlograms.  This
vignette reproduces most of the figures in @friendly2002corrgrams.


## Setup

```{r setup, results="hide"}
library("knitr")
opts_chunk$set(fig.align="center", fig.width=6, fig.height=6)
options(width=90)
```


## Figure 2

```{r fig2}
library("corrgram")
vars2 <- c("Assists","Atbat","Errors","Hits","Homer","logSal",
           "Putouts","RBI","Runs","Walks","Years")
corrgram(baseball[,vars2], order=TRUE,
         main="Baseball data PC2/PC1 order",
         lower.panel=panel.shade, upper.panel=panel.pie,
         diag.panel=panel.minmax, text.panel=panel.txt)
```


## Figure 3

```{r fig3}
baseball.cor <- cor(baseball[,vars2], use='pair')
baseball.eig <- eigen(baseball.cor)$vectors[,1:2]
e1 <- baseball.eig[,1]
e2 <- baseball.eig[,2]
plot(e1,e2,col='white', xlim=range(e1,e2), ylim=range(e1,e2))
text(e1,e2, rownames(baseball.cor), cex=1)
arrows(0, 0, e1, e2, cex=0.5, col="red", length=0.1)
```


## Figure 4a, 4b

```{r fig4}
corrgram(baseball[,vars2], main="Baseball data (alphabetic order)")

corrgram(baseball[,vars2], order=TRUE,
         main="Baseball data (PC order)",
         panel=panel.shade, text.panel=panel.txt)
```


## Figure 5
```{r fig5}
corrgram(baseball, order=TRUE, main="Baseball data (PC order)")
```


## Figure 6. Arrangement is slightly different from Friendly.

```{r fig6}
corrgram(auto, order=TRUE, main="Auto data (PC order)")
```


## Figure 7.

```{r fig7}
rinv <- function(r){
  # r is a correlation matrix
  # calculate r inverse and scale to correlation matrix
  # Derived from Michael Friendly's SAS code

  ri <- solve(r)
  s <- diag(ri)
  s <- diag(sqrt(1/s))
  ri <- s %*% ri %*% s
  n <- nrow(ri)
  ri <- ri * (2*rep(1,n) - matrix(1, n, n))
  diag(ri) <- 1  # Should already be 1, but could be 1 + epsilon
  colnames(ri) <- rownames(ri) <- rownames(r)
  return(ri)
}

vars7 <- c("Years", "logSal", "Homer", "Putouts", "RBI", "Walks",
           "Runs", "Hits", "Atbat", "Errors", "Assists")
cb <- cor(baseball[,vars7], use="pair")
corrgram(-rinv(cb), main=expression(paste("Baseball data ", R^-1)))

```


## Figure 8

```{r fig8}
require(Matrix) # For block diagonal function

partial <- function(r, xvar){
  # r is a correlation matrix
  # Calculate partial correlation of y|x
  yvar <- setdiff(colnames(r), xvar)
  ri <- r[yvar,yvar] - r[yvar,xvar] %*% solve(r[xvar,xvar]) %*% r[xvar,yvar]
  s <- diag(ri)
  s <- diag(sqrt(1/s))
  ri <- s %*% ri %*% s
  ri <- as.matrix(Matrix::bdiag(ri, r[xvar, xvar]))
  diag(ri) <- 1  # Should already be 1, but could be 1 + epsilon
  colnames(ri) <- rownames(ri) <- c(yvar, xvar)
  return(ri)
}

vars8a <- c("Gratio", "Rep78", "Rep77", "Hroom", "Trunk", "Rseat",
            "Length", "Weight", "Displa", "Turn")
vars8b <- c("MPG", "Price")
vars8 <- c(vars8a, vars8b)
auto.cor <- cor(auto[, vars8], use="pair")
auto.par <- partial(auto.cor, vars8b)
corrgram(auto.par, lower.panel=panel.pie, upper.panel=panel.pie,
         main="Auto data, partialing out Price,MPG")
```


## Figure 11

```{r fig11}
corrgram(baseball[,vars2], order=TRUE,
         main="Baseball correlation ellipses",
         panel=panel.ellipse,
         text.panel=panel.txt, diag.panel=panel.minmax)
```


## Session information
```{r session}
sessionInfo()
```

## References
